### 2018-10-28 paper/complexity.tex
update

Summing up  formula 
(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  = 
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an error-free
SUT $Q$ against reference process $\pmax$. This maximum is met when $Q$ equals $\Sigma^*$.

@@

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ and study the potential increase of test executions if
the behaviour of $Z$ is changed to $(Z/\ol s)\intchoice \Stop$ after some specific 
trace $\ol s\in\trc(Z)$ with $\minhits(P/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
an can be specified by\footnote{We use the following prefix operators
for given   traces $s_1,s_2\in\Sigma^*$: the Boolean 
expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.}
\begin{eqnarray}
Z({\ol s}) & = & Z({\ol s},\varepsilon)
\\
\label{eq:zstop}
Z(\ol s,s) & = & (s = \ol s) \& ((Z/s)\intchoice \Stop)
\\ & & \Extchoice    \nonumber
\\  
\label{eq:zstoppref}
& & (s < \ol s \wedge \minhits(P/s) = \varnothing) \& 
\big(  (e:[P/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
\\ & & \Extchoice   \nonumber
\\ 
\label{eq:zcont}
& &           (s < \ol s\wedge \minhits(P/s) \neq \varnothing ) \& 
\\ & & \nonumber
\big(\Intchoice_{A\in\minaccs(P/s)\cup \{ [P/s]^0 \}}
e : A \then Z(\ol s, s.e)          \big)
\\ & & \Extchoice   \nonumber
\\
\label{eq:zlikep}
& & (s \not\le \ol s) \& (P/s)
\end{eqnarray}
%
If $Z(\ol s,s)$ has executed trace $\ol s$, its branch (\ref{eq:zstop}) is entered, 
so the process 
nondeterministically decides whether to continue like $Z/s$ or to stop. As a consequence,
$Z(\ol s)/\ol s = Z(\ol s,\ol s)$ only has $\varnothing$ as minimal acceptance, and its
  collection of minimal hitting sets is empty.
 While $Z(\ol s)$ still executes a true prefix
of $\ol s$, it behaves exactly as $Z$, executing either branch (\ref{eq:zstoppref}) or
(\ref{eq:zcont}): the latter branch is taken if $P/s$ has at least one minimal hitting set,
so that it cannot refuse all elements of $[P/s]^0$. In such a case, $P/s$ can be
represented in a normalised form, where an initial event is selected from an internal
choice over all minimal acceptances and over $[P/s]^0$ (the choice $[P/s]^0$ is needed
because the union over all minimal acceptances is not necessarily identical 
to $[P/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
The former branch is taken if $P/s$ can block all of its initials.
As a consequence, in both branches, $Z(\ol s,s)$ 
has the same nondeterministic behaviour and the same initials as $P/s$, as long as
a prefix of $\ol s$ is executed. As soon as $Z(\ol s)$ has run through a trace $s$ that is no longer a prefix of $\ol s$, 
the process continues exactly as $Z$ would do after having run through $s$. This situation
is captured by branch (\ref{eq:zlikep}).
Summarising, $Z(\ol s)$ behaves like $P$ for all but the
one trace   $\ol s$.

@@

\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has $\card{\minhits(Z/\ol s)}$ fewer executions than $(Z(\ol s)\parallel[\Sigma] U_F(j))$, because 
$Z(\ol s,\ol s)$ has an empty set of hitting sets. As a consequence, test case
${\ol U}_F(\#\ol s)$ cannot enter branch  
\end{enumerate}

>>>>

Summing up  formula 
(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  = 
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an error-free
SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.

@@

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ and study the potential increase of test executions if
the behaviour of $Z$ is changed to $(Z/\ol s)\intchoice \Stop$ after some specific 
trace $\ol s\in\trc(Z)$ with $\minhits(Z/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
an can be specified by\footnote{We use the following prefix operators
for given   traces $s_1,s_2\in\Sigma^*$: the Boolean 
expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.}
\begin{eqnarray}
Z({\ol s}) & = & Z({\ol s},\varepsilon)
\\
\label{eq:zstop}
Z(\ol s,s) & = & (s = \ol s) \& ((Z/s)\intchoice \Stop)
\\ & & \Extchoice    \nonumber
\\  
\label{eq:zstoppref}
& & (s < \ol s \wedge \minhits(Z/s) = \varnothing) \& 
\big(  (e:[Z/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
\\ & & \Extchoice   \nonumber
\\ 
\label{eq:zcont}
& &           (s < \ol s\wedge \minhits(Z/s) \neq \varnothing ) \& 
\\ & & \nonumber
\big(\Intchoice_{A\in\minaccs(Z/s)\cup \{ [Z/s]^0 \}}
e : A \then Z(\ol s, s.e)          \big)
\\ & & \Extchoice   \nonumber
\\
\label{eq:zlikep}
& & (s \not\le \ol s) \& (Z/s)
\end{eqnarray}
%
If $Z(\ol s,s)$ has executed trace $\ol s$, its branch (\ref{eq:zstop}) is entered, 
so the process 
nondeterministically decides whether to continue like $Z/s$ or to stop. As a consequence,
$Z(\ol s)/\ol s = Z(\ol s,\ol s)$ only has $\varnothing$ as minimal acceptance, and its
  collection of minimal hitting sets is empty.
 While $Z(\ol s)$ still executes a true prefix
of $\ol s$, it behaves exactly as $Z$, executing either branch (\ref{eq:zstoppref}) or
(\ref{eq:zcont}): the latter branch is taken if $Z/s$ has at least one minimal hitting set,
so that it can never refuse all elements of $[Z/s]^0$. In such a case, $Z/s$ can be
represented in a normalised form, where an initial event is selected from an internal
choice over all minimal acceptances and over $[Z/s]^0$ (the choice $[Z/s]^0$ is needed
because the union over all minimal acceptances is not necessarily identical 
to $[Z/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
The former branch is taken if $Z/s$ can block all of its initials.
As a consequence, in both branches, $Z(\ol s,s)$ 
has the same nondeterministic behaviour and the same initials as $Z/s$, as long as
a prefix of $\ol s$ is executed. As soon as $Z(\ol s)$ has run through a trace $s$ that is no longer a prefix of $\ol s$, 
the process continues exactly as $Z$ would do after having run through $s$. This situation
is captured by branch (\ref{eq:zlikep}).
Summarising, $Z(\ol s)$ behaves like $P$ for all but the
one trace   $\ol s$.

@@

\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has 
$(\card{\minhits(Z/\ol s)} - 1)$ fewer executions than $(Z\parallel[\Sigma] U_F(j))$, because 
$Z(\ol s,\ol s)$ has an empty set of hitting sets, since 
$Z(\ol s,\ol s) = (Z/\ol s)\intchoice\Stop$. As a consequence, test case
${\ol U}_F(\#\ol s)$ can only enter branches 
(\ref{eq:ufa})  or (\ref{eq:zcont}) after having run through trace
$\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.

\item For $j > \#\ol s$, one execution is won by entering 
branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
\end{enumerate}
%
Summing up over all test cases for $j=0,\dots,pq-1$, the difference in test executions generated
from $Z(\ol s)$ and those generated from  $Z$ is
\begin{equation}
 pq - \#\ol s       - \card{\minhits(Z/\ol s)}.
\end{equation}
%
This expression is maximal for $\card{\minhits(Z/\ol s)} = 1$ (recall that
the definition of $Z(\ol s)$ requires $\card{\minhits(Z/\ol s)} > 0$). 
The number of hitting sets in $\minhits(Z/\ol s)$ is one if and only if the
minimal acceptances of $Z/\ol s$ have exactly one element in their common intersection.
The maximal length of 
$\ol s$ where the expression remains strictly positive is then $pq - 2$.

The construction of $Z(\ol s)$ can be generalised by using $k$ traces 
${\ol s}_i, \ i = 1\dots,k$ of the same length $\ol j$, 
after that the modified process behaves as 
$(Z/{\ol s}_i)\intchoice \Stop$. Generalising the case distinction above, 
\begin{enumerate}
\item the number of test executions remains unchanged for $j < \ol j$,
\item the number of   executions lost for $j = \ol j$ is at least 
$k$ (assuming that for all $i$, $\card{\minhits(Z/{\ol s}_i)} = 1$), and
\item the number of executions won for $j > \ol j$ is $k$.
\end{enumerate}
%
The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ is bounded by the maximal number
of traces that can be executed by $(Z\parallel[\Sigma] U_F(j))$, which is bounded by
$\card{\Sigma}^{\ol j}$ (this bound is reached, for example, 
for process $\pmax$ discussed above). If $k$ equals this maximal number, there is no need
to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.

Summing up over all test cases for such a $k \le \card{\Sigma}^{\ol j}$, this results in 
an upper bound of
\begin{equation}
 \card{\Sigma}^{\ol j}(pq - \ol j      - 1)
\end{equation}
additional test executions in comparison to the number of executions needed for all tests
of $Z$. Now we can apply this construction to all traces lengths from $0$ to $pq-2$ (we have 
explained above that test executions are lost for trace lengths greater than $pq-2$). This results in an upper bound of
\begin{equation}
\sum_{j=0}^{pq-2}  n^j(pq - j      - 1) =  
\frac{n^{pq} + pq - npq -1}{(n - 1)^2}
\quad\text{with $n = \card{\Sigma}$}
\end{equation}
additional test executions that can be gained by applying the construction technique
explained above for $Z(\ol s)$, extended to all traces of length less or equal to $pq-2$.
%
Summarising this result and the one previously obtained in (\ref{eq:maxexec}),
the following complexity bounds are obtained.
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} has the following bounds.
%
\begin{enumerate}
\item If $pq \le \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
test suite  
requires at most 
%
\begin{equation*}
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
\end{equation*}
%
test executions. This upper bound is reached for process $\pmax$ specified above.

\item If $pq > \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
test suite  
requires less than 
\begin{equation*}
pq\frac{1-n^{pq}}{1-n} +
\frac{n^{pq} + pq - npq -1}{(n - 1)^2} \qquad\text{@todo}
\end{equation*}
executions.
\end{enumerate}
\xbox
\end{theorem}
